### 2019-03-21 report/paper.tex
Fill the report.

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{matrix}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{COMP 6721: Project Report}
\author{Martin Khannouz}
%
\institute{40093517 \email{martin.khannouz@gmail.com}}
%
\maketitle              % typeset the header of the contribution
%
\section*{Introduction}
This project report describes how the first
project of \textit{Introduction to Artificial
Intelligence} were developed. The first section
covers the environment and the tools used for the
project. The heuristic section describes each
heuristic and how they were improved. The
difficulties section summerizes the difficulties
encountered during the project and how they were
overcome.

\section{Settings and Structures}
\subsection{Settings}
The developing part of this project was realized
on an Archlinux distribution with Python~3.7.2.
Vim with a Jedi plugin was used for the coding
part and a private Github repository was set to
keep track of the changes.
The \textit{orwell} machine was used before each
deliverable to ensure the project worked with the
software version on Concordia's computers. Pytest
was used to guarantee no regression has been made.

\subsection{stucture}
The project is structured in classes with
the main one being the engine. This class is an
instance of the game that implements the rules.
To ensure this class operates properly, unit tests
were implemented before the class methods, based on
the use cases given by the teacher.
This class was optimized as much as possible to
guarantee the fastest execution time for minimax
player.

The project also contains three type of players. A
\textit{Human} player that parse the input from the command
line. A \textit{MinMax} player that implement tweaked
minimax version with alpha-beta pruning. The
\textit{MonteCarlo} player which was supposed to implement
a Monte Carlo Tree Search even though it was a
complete failure due to the poor understanding of
this algorithm by the author of the project. Finally, a random player was
developed during the early stage of the project.
This "algorithm" served as control test for other
player types, because none of them should lost
against such chaotic player\footnote{Based on my
statistics, even a 5-years old kid should get a
win rate of 50\%.}.

Automatic players (\textit{MinMax} and
\textit{MonteCarlo}) rely on heuristic to complete
their search. All heuristic functions are grouped
in one file: \textit{heuristic.py}.
Finally, the class \textit{Move} represents a move
that could be played.

In addition to these classes, additional functions
were implemented. One function used genetic
algorithms to optimize the \textit{Vspace}
heuristic, but it was too slow and just gave hint
on how to enhance the heuristic rather than a
accurate weights.  A win rate function was also
implemented to compare combinations of player and
heuristic against each other.

\section{Heuristics}
Five heuristics were developed in this project
even though only two were made for competitive
use. One heuristic is the naive function for the
second deliverable. Another is a random heuristic
that systematicaly return a new random number. It
was used to spot runtime errors in \textit{MinMax} and
\textit{MonteCarlo} algorithms. The \textit{Basic}
heuristics only check for winning situation and
was used for basic controls.
The two other heuristics are described in
subsection of their own.

\subsection{Convolution}
The convolution utilize the \textit{convolve2d} function
from Scipy to estimate the value of a current
state.
The game board is split into two matrices: dot and
color. Each matrix contains -1, 0, or 1. A zero
indicates an empty space while the two other values
represent either a filled dot and an empty dot, or
a white cell and a red cell.
Applying \textit{convolve2d} on these matrices
produces new matrices with values that range from
-4 to 4. An absolute value of four indicates a row
of four and therefore, a winning position.
The heuristic gathers all intermediate values into
an histogram (using the \textit{histogram}
function of numpy) then give a weight to each
value.

Figure~\ref{fig:board-example} shows an example of
the board. Figure~\ref{fig:dot-representation}
focus on the dot player and show the
transformation applied to the board for this
player. Filled dots are set to 1 and circles to
-1. Zeros are not represented for more clarity.
The heuristic applies four convolutions: one for
the rows, one for the columns and one for each
diagonal. These convolution generate four smaller
matrices showed in
Figure~\ref{fig:convolution-results}.  Then, the
matrices are flatened, concateneted and the
absolute value is applied to every cell. Finally
the histogram is built out of this matrix and each
element is multiplied by a weight before being
summed.  The final
value of the board is the difference between the
two player values.

This approach is strong for many reason. It is
faster than calling python code because Scipy and
Numpy rely on C writen functions. Uncomplete rows
that contain zeros are more valued than uncomplete
rows that contain opposite values. 
\begin{figure}[ht]
		\center
		\begin{tikzpicture}

				\node [draw, fill=red!20, minimum width=0.5cm, minimum height=0.5cm] at (0, 0) {$\bullet$};
				\node [draw, fill=red!20, minimum width=0.5cm, minimum height=0.5cm] at (0.5, 0) {$\circ$};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.0, 0) {};
				\node [draw, fill=red!20, minimum width=0.5cm, minimum height=0.5cm] at (1.5, 0) {$\bullet$};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (2.0, 0) {};

				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0, 0.5) {$\circ$};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0.5, 0.5) {$\bullet$};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.0, 0.5) {};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.5, 0.5) {$\circ$};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (2.0, 0.5) {};

				\node [draw, fill=red!20, minimum width=0.5cm, minimum height=0.5cm] at (0, 1) {$\circ$};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0.5, 1) {$\circ$};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.0, 1) {};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.5, 1) {};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (2.0, 1) {};

				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0, 1.5) {$\bullet$};
				\node [draw, fill=red!20, minimum width=0.5cm, minimum height=0.5cm] at (0.5, 1.5) {$\bullet$};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.0, 1.5) {};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.5, 1.5) {};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (2.0, 1.5) {};

		\end{tikzpicture}
		\caption{An example of a board game. Cards are not represented.}
		\label{fig:board-example}
\end{figure}
\begin{figure}[ht]
		\center
		\begin{tikzpicture}

				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0, 0) {$\bullet$};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0.5, 0) {$\circ$};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.0, 0) {};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.5, 0) {$\bullet$};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (2.0, 0) {};

				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0, 0.5) {$\circ$};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0.5, 0.5) {$\bullet$};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.0, 0.5) {};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.5, 0.5) {$\circ$};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (2.0, 0.5) {};

				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0, 1) {$\circ$};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0.5, 1) {$\circ$};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.0, 1) {};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.5, 1) {};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] (A) at (2.0, 1) {};

				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0, 1.5) {$\bullet$};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0.5, 1.5) {$\bullet$};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.0, 1.5) {};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.5, 1.5) {};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (2.0, 1.5) {};

				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (5, 0) {1};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (5.5, 0) {-1};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (6.0, 0) {};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (6.5, 0) {1};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (7.0, 0) {};

				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (5, 0.5) {-1};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (5.5, 0.5) {1};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (6.0, 0.5) {};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (6.5, 0.5) {-1};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (7.0, 0.5) {};

				\node [draw, minimum width=0.5cm, minimum height=0.5cm] (B) at (5, 1) {-1};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (5.5, 1) {-1};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (6.0, 1) {};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (6.5, 1) {};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (7.0, 1) {};

				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (5, 1.5) {1};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (5.5, 1.5) {1};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (6.0, 1.5) {};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (6.5, 1.5) {};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (7.0, 1.5) {};

				\draw [->,thick] (A) -- (B);
		\end{tikzpicture}
		\caption{Example of transformation for the dot player based on the board in Figure~\ref{fig:board-example}.}
		\label{fig:dot-representation}
\end{figure}
\begin{figure}[ht]
		\center
		\begin{tikzpicture}

				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0, 0) {1};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0.5, 0) {0};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0, 0.5) {-1};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0.5, 0.5) {0};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0, 1) {-2};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0.5, 1) {-1};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0, 1.5) {2};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (0.5, 1.5) {1};

				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.5, 0) {0};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (2.0, 0) {0};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (2.5, 0) {0};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (3.0, 0) {0};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] (C1) at (3.5, 0) {0};

				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (1.5, 1) {1};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (2.0, 1) {0};

				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (3.0, 1) {2};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] (C0) at (3.5, 1) {-1};

				\node [draw, minimum width=0.5cm, minimum height=0.5cm] (H0) at (7.0, 0) {8};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (7.5, 0) {6};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (8.0, 0) {3};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (8.5, 0) {0};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] at (9.0, 0) {0};

				\node [draw, minimum width=0.5cm, minimum height=0.5cm] (W0) at (7.0, 0.75) {0};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] (W1) at (7.5, 0.75) {1};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] (W2) at (8.0, 0.75) {8};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] (W3) at (8.53, 0.75) {16};
				\node [draw, minimum width=0.5cm, minimum height=0.5cm] (W4) at (9.15, 0.75) {$10^6$};
				\node at (10.5, 0.75) {Weights};
				\node at (10.5, 0) {Histogram};
				\node at (10.5, -0.5) {Indexes};

				\node at (7.0, -0.5) {0};
				\node at (7.5, -0.5) {1};
				\node at (8.0, -0.5) {2};
				\node at (8.5, -0.5) {3};
				\node at (9.0, -0.5) {4};

				\node (ID0) at (7.0, 2) {$\underbrace{0\times8} + $};
				\node (ID1) at (8.0, 2) {$\underbrace{1\times6} +$};
				\node (ID2) at (9.0, 2) {$\underbrace{8\times3} +$};
				\node (ID3) at (10.15, 2) {$\underbrace{16\times0} +$};
				\node (ID4) at (11.35, 2) {$\underbrace{10^6 \times 0}$};
				\node (ID5) at (12.5, 2.10) {$ = 30$};

				\draw [->,thick] (W0) -- (ID0);
				\draw [->,thick] (W1) -- (ID1);
				\draw [->,thick] (W2) -- (ID2);
				\draw [->,thick] (W3) -- (ID3);
				\draw [->,thick] (W4) -- (ID4);

				\draw [->,thick] (C0) -- (H0);
				\draw [->,thick] (C1) -- (H0);
		\end{tikzpicture}
		\caption{The four matrices obtained after applying the four masks on the matrix that represent the dot, in Figure~\ref{fig:dot-representation}.
						This Figure also represents the conversion to an histogram and how this histogram is used to find the heuristic value for the dot player.}
		\label{fig:convolution-results}
\end{figure}

\subsection{Vspace}
The Vspace heuristic tries to give value to empty
cells in addition to counting rows of similar
symbols. This heuristic count the number of lines
and their sizes.
\section{Difficulties}
\begin{itemize}
		\item Slowness of python, (numpy first)
		\item Version changing between Python 3.7.2 and 3.5.1 (serialization of regex)
\end{itemize}
\section{Analyze}
\begin{itemize}
		\item How is the pruning.
		\item How are the performance at the tournament.
\end{itemize}
\section{A Subsection Sample}
Please note that the first paragraph of a section or subsection is
not indented. The first paragraph that follows a table, figure,
equation etc. does not need an indent, either.

Subsequent paragraphs, however, are indented.

\subsubsection{Sample Heading (Third Level)} Only two levels of
headings should be numbered. Lower level headings remain unnumbered;
they are formatted as run-in headings.

\paragraph{Sample Heading (Fourth Level)}
The contribution should contain no more than four levels of
headings. Table~\ref{tab1} gives a summary of all heading levels.

\begin{table}
\caption{Table captions should be placed above the
tables.}\label{tab1}
\begin{tabular}{|l|l|l|}
\hline
Heading level &  Example & Font size and style\\
\hline
Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
\hline
\end{tabular}
\end{table}

\noindent Displayed equations are centered and set on a separate
line.
\begin{equation}
x + y = z
\end{equation}
Please try to avoid rasterized images for line-art diagrams and
schemas. Whenever possible, use vector graphics instead (see
Fig.~\ref{fig1}).

\begin{figure}
\includegraphics[width=\textwidth]{fig1.eps}
\caption{A figure caption is always placed below the illustration.
Please note that short captions are centered, while long ones are
justified by the macro package automatically.} \label{fig1}
\end{figure}

\begin{theorem}
This is a sample theorem. The run-in heading is set in bold, while
the following text appears in italics. Definitions, lemmas,
propositions, and corollaries are styled the same way.
\end{theorem}
%
% the environments 'definition', 'lemma', 'proposition', 'corollary',
% 'remark', and 'example' are defined in the LLNCS documentclass as well.
%
\begin{proof}
Proofs, examples, and remarks have the initial word in italics,
while the following text appears in normal font.
\end{proof}
For citations of references, we prefer the use of square brackets
and consecutive numbers. Citations using labels or the author/year
convention are also acceptable. The following bibliography provides
a sample reference list with entries for journal
articles~\cite{ref1}, an LNCS chapter~\cite{ref2}, a
book~\cite{ref1}, proceedings without editors~\cite{ref1},
and a homepage~\cite{ref_2}. Multiple citations are grouped
\cite{ref1,ref2}.
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%
\bibliographystyle{splncs04}
\bibliography{paper}
\end{document}

% vim: tw=50 ts=2